---
title: "immunogenicity benchmark"
author: "Guangshuai Wang"
date: "`r Sys.Date()`"
output:   
  rmdformats::readthedown:
    highlight: kate
    lightbox: false
    toc_depth: 3
    mathjax: true
---

```{r , include=FALSE}
options(max.print = "120")
knitr::opts_chunk$set(echo = TRUE, comment = "#>", eval = TRUE, collapse = TRUE,cache = FALSE,warning = FALSE)
knitr::opts_knit$set(width = 120)
```

```{r , include=FALSE}
library(tidyverse)
library(reticulate)
library(DT)
use_python(Sys.which("python"))
```

```{python, include = FALSE}
import pandas as pd
import matplotlib.pyplot as plt
import matplotlib as mpl
mpl.rcParams['pdf.fonttype'] = 42
mpl.rcParams['ps.fonttype'] = 42
plt.rcParams.update({'font.family':'Arial'})
import seaborn as sns
import numpy as np
from sklearn.metrics import confusion_matrix,accuracy_score
from sklearn.metrics import roc_curve,auc,confusion_matrix,precision_score,recall_score,precision_recall_curve
import matplotlib as mpl
import sys
sys.path.append("../Python")
from Delongtest import DelongTest
```

we compare with another immunogenicity preditced tools

## Benchmark in immunoigenicity

### Data proprecessing

form nature data, sub k-mer method

```{python}
#python
peptide = pd.read_csv("../data/nature_peptide.csv")
HLA = pd.read_csv("../data/nature_HLA.csv")
peptide = peptide.drop_duplicates()
peptide["Immunogenicity"].value_counts()
peptide["Len"] = peptide["Peptide"].map(len)
peptide[peptide["Len"]>=15]["Peptide"].to_csv("../data/pep_only.txt",header = None,index = None,sep = " ")#chain indexing
IEDB_pep = peptide[peptide["Len"]>=15].reset_index(drop = True)
IEDB_pep["num"] = [x+1 for x in range(len(IEDB_pep))]#used to predict by IEDB tools
pep_ID = ["Pep{}".format(x+1) for x in range(len(peptide))]
peptide["pep_ID"] = pep_ID
fig,ax = plt.subplots()
sns.countplot(data = peptide,x = "Len")
plt.show()
```

```{python}
#python
ax = sns.countplot(data=peptide, x="Immunogenicity")
#ax.bar_label(ax.containers[0])
ax.text(-0.1, 860, "{:.3f}".format(851/931))
ax.text(0.9, 90, "{:.3f}".format(80/931))
plt.savefig("../figure/nature_immuno_rate.pdf",dpi = 300,transparent=True)
plt.show()
```

```{python}
#python
Data = pd.merge(HLA,peptide)
IMM_train_data_allele = pd.read_csv("../data/IEDB_MHCII_immuno_allele.csv",header = None,names = ["HLA"])
Data = Data[Data["HLA"].isin(IMM_train_data_allele["HLA"])]
pseudo_seq = pd.read_table("../data/pseudosequence.2016.all.X.dat",header = None,names = ["HLA","sequence"])
Data1 = pd.merge(Data,pseudo_seq)
Data1["length"] = Data1["Peptide"].map(len)
Data1 = Data1[Data1["length"]>=13]
Data1 = Data1.reset_index(drop = True)
#15mer
def mer15(Pep):
    P = []
    Length = 15
    for i in range(len(Pep) - Length +1):
        pep = Pep[i:i+Length]
        P.append(pep)
    return P
DF15 = pd.DataFrame()
for i in range(len(Data1)):
    pep = mer15(Data1["Peptide"][i])
    df = pd.DataFrame(pep)
    df["pep_ID"] = Data1["pep_ID"][i]
    df["HLA"] = Data1["HLA"][i]
    DF15 = pd.concat([DF15,df])
DF15.columns = ["pep","pep_ID","HLA"]
benchmark_data = pd.merge(DF15,pseudo_seq)
benchmark_data.to_csv("../data/15mer_data.csv")
benchmark_data["pep"].to_csv("../data/15mer_pep_only.txt",header = None,index = None,sep = " ")
#kmer
def kmer(Pep,sa,end):
    P = []
    R = range(sa,end+1)
    for i in R:
        Length = i
        for i in range(len(Pep) - Length +1):
            pep = Pep[i:i+Length]
            P.append(pep)
    return P
DF_k = pd.DataFrame()
for i in range(len(Data1)):
    pep = kmer(Data1["Peptide"][i],13,21)
    df = pd.DataFrame(pep)
    df["pep_ID"] = Data1["pep_ID"][i]
    df["HLA"] = Data1["HLA"][i]
    DF_k = pd.concat([DF_k,df])
DF_k.columns = ["pep","pep_ID","HLA"]
benchmark_data = pd.merge(DF_k,pseudo_seq)
benchmark_data.to_csv("../data/kmer_data.csv")
```

### run

#### IEDB

[IEDB onoline prediction](http://tools.iedb.org/CD4episcore/)

#### Repitope

[Repitope](https://github.com/masato-ogishi/Repitope)

```{r,eval=FALSE}
#PBS
#R
library(tidyverse)
library(data.table)
library(Repitope)
options(java.parameters="-Xmx60G")

fragLibDT <- fst::read_fst("~/Project/MHCII/Repitope/FragmentLibrary_TCRSet_Public_RepitopeV3.fst", as.data.table=T)
featureDF_MHCII <- fst::read_fst("~/Project/MHCII/Repitope/FeatureDF_MHCII_Weighted.10000_RepitopeV3.fst", as.data.table=T)

data <- read_csv("~/Project/MHCII/Repitope/15mer_data.csv")
peptideSet_of_interest <- data$pep
peptideSet=peptideSet_of_interest
len = sapply(data$pep, nchar)
table(len)

res_MHCII <- EpitopePrioritization(
  featureDF=featureDF_MHCII[Peptide%in%MHCII_Human$Peptide,], 
  metadataDF=MHCII_Human[,.(Peptide,Immunogenicity)],
  peptideSet=peptideSet_of_interest,
  peptideLengthSet=11:30,
  fragLib=fragLibDT,
  aaIndexIDSet="all",
  fragLenSet=3:11,
  fragDepth=10000,
  fragLibType="Weighted",
  featureSet=MHCII_Human_MinimumFeatureSet,
  seedSet=1:1,
  coreN=1,
  outDir="~/Project/MHCII/Repitope/pre_result"  ## Intermediate and final output files will be stored under this directory
)
```

#### netMHCIIpan BA&EL

This part code is similar to (添加饮用), only chage the input and output file, so we don't add code in this part.

#### TLimmuno2

This part code is similar to (添加饮用), only chage the input and output file, so we don't add code in this part.

### Result

```{python 15mer}
#python
def IMM_process(Data,peptide):
    sam_pep_ID = Data["pep_ID"].unique()
    res = []
    p_ID = []
    Data["length"] = Data["pep"].map(len)
    for a in sam_pep_ID:
        Mean = []    
        ID = Data[Data["pep_ID"] == a]
        for i in ID["HLA"].unique():
            Data_HLA = ID[ID["HLA"] == i]
            prediction = []
            for l in Data_HLA["length"].unique():
                Data_len = Data_HLA[Data_HLA["length"] == l]
                pre = Data_len["prediction"].max()
                prediction.append(pre)
            x = np.max(prediction)
            Mean.append(x)
        res.append(max(Mean))
        p_ID.append(a)
    result = pd.DataFrame({"pep_ID":p_ID,"prediction":res})
    result = result.merge(peptide,how = "inner",on = "pep_ID" )
    return result
  
def auc_pr(data,true_label,prediction,rank = False):
    if rank:
        fpr,tpr,_ = roc_curve(data[true_label],1-data[prediction])
        precision,recall,_ = precision_recall_curve(data[true_label],1-data[prediction])
    else:
        fpr,tpr,_ = roc_curve(data[true_label],data[prediction])
        precision,recall,_ = precision_recall_curve(data[true_label],data[prediction])
        
    AUC = auc(fpr,tpr)
    PR = auc(recall,precision)
    
    return fpr,tpr,recall,precision,AUC,PR
  
result = pd.read_csv("../data/15mer_result.csv")
result_15mer = IMM_process(result,peptide)
mer15_result = auc_pr(result_15mer,"Immunogenicity","prediction")
print("AUC:{}".format(mer15_result[4]))
print("PR_AUC:{}".format(mer15_result[5]))
```

```{python kmer}
#python
result = pd.read_csv("../data/kmer_result.csv")
result_kmer = IMM_process(result,peptide)
result_kmer = result_kmer[result_kmer["pep_ID"].isin(result_15mer["pep_ID"])]
kmer_result = auc_pr(result_kmer,"Immunogenicity","prediction")
print("AUC:{}".format(kmer_result[4]))
print("PR_AUC:{}".format(kmer_result[5]))
```

```{python IEDB_online_tool}
#python
IEDB_res = pd.read_csv("../data/IEDB_CD4_tools_manually.csv")
Imm_score = []
pro_num = []
for i in IEDB_res["Protein Number"].unique():
    y = IEDB_res[IEDB_res["Protein Number"] == i]
    imm_score = y["Immunogenicity Score"].mean()
    pro_num.append(i)
    Imm_score.append(imm_score)
IEDB_result = pd.DataFrame({"num":pro_num,"score":Imm_score})
IEDB_result = pd.merge(IEDB_result,IEDB_pep)
IEDB_result = IEDB_result[IEDB_result["Peptide"].isin(result_15mer["Peptide"])]
IEDB_score = auc_pr(IEDB_result,"Immunogenicity","score",rank = True)
print("AUC:{}".format(IEDB_score[4]))
print("PR_AUC:{}".format(IEDB_score[5]))
```


```{python Repitope}
#python
Repitope_result = pd.read_csv("../data/repitope_15mer_result.csv")
Repitope = pd.merge(Repitope_result,DF15,left_on = "Peptide",right_on = "pep")
sam_pep_ID = Repitope["pep_ID"].unique()
res = []
p_ID = []
for a in sam_pep_ID:
    Mean = []
    ID = Repitope[Repitope["pep_ID"] == a]
    for i in ID["HLA"].unique():
        x = ID[ID["HLA"] == i]
        x_mean = x["ImmunogenicityScore"].mean()
        Mean.append(x_mean)
    res.append(max(Mean))
    p_ID.append(a)
dic = {"pep_ID":p_ID,"ImmunogenicityScore":res}
Repitope_pre = pd.DataFrame(dic)
Repitope_result = Repitope_pre.merge(peptide,how = "inner",on = "pep_ID" )
Repitope_score = auc_pr(Repitope_result,"Immunogenicity","ImmunogenicityScore")
print("AUC:{}".format(Repitope_score[4]))
print("PR_AUC:{}".format(Repitope_score[5]))
```

```{python netMHCIIpan}
#python
def netMHCIIpan_process(result,allele,ori_sorted,peptide):
    
    #combine two type allele
    result = pd.merge(result,allele,left_on="allele",right_on="Allele Name")
    result_sorted = result.sort_values(by=["peptide","HLA"]).reset_index(drop = True)
    #combine with ori data
    result_combined = pd.merge(result_sorted,ori_sorted[["pep","pep_ID"]],left_on = "peptide",right_on ="pep")
    #result_combined["pep_length"] = result_combined["peptide"].map(len)
    sam_pep_ID = result_combined["pep_ID"].unique()
    res = []
    p_ID = []
    for a in sam_pep_ID:
        Mean = []
        ID = result_combined[result_combined["pep_ID"] == a]
        #get Mean of all splited pep and Max of all HLA
        for i in ID["Allele Name"].unique():
            Data_HLA = ID[ID["Allele Name"] == i]
            prediction = []
            for l in Data_HLA["pep_length"].unique():
                Data_len = Data_HLA[Data_HLA["pep_length"] == l]
                pre = Data_len["percentile_rank"].min()
                prediction.append(pre)
            x = np.min(prediction)
        res.append(np.min(x))
        p_ID.append(a) 
    final_result = pd.DataFrame({"pep_ID":p_ID,"prediction":res})
    final_result = final_result.merge(peptide,how = "inner",on = "pep_ID" )
    return final_result 
```

```{python}
DF15_sorted = DF15.sort_values(by=["pep","HLA"]).reset_index(drop = True)
allele = pd.read_csv("/Users/wangguangshuai/Data/MHCII/Data/IMM_model_benchmark/allele.csv")
netMHCIIpan_ba = pd.read_csv("/Users/wangguangshuai/Data/MHCII/Data/IMM_model_benchmark/netMHCIIpan_15_mer_ba.csv")
netMHCIIpan_ba["pep_length"] = netMHCIIpan_ba["peptide"].map(len)
net_15mer_ba = netMHCIIpan_process(netMHCIIpan_ba,allele,DF15_sorted,peptide)
mer15_ba_result = auc_pr(net_15mer_ba,"Immunogenicity","prediction",rank = True)
print("AUC:{}".format(mer15_ba_result[4]))
print("PR_AUC:{}".format(mer15_ba_result[5]))
```

```{python}
netMHCIIpan_el = pd.read_csv("/Users/wangguangshuai/Data/MHCII/Data/IMM_model_benchmark/netMHCIIpan_15_mer_el.csv")
netMHCIIpan_el["pep_length"] = netMHCIIpan_el["peptide"].map(len)
net_15mer_el = netMHCIIpan_process(netMHCIIpan_el,allele,DF15_sorted,peptide)
mer15_el_result = auc_pr(net_15mer_el,"Immunogenicity","prediction",rank = True)
print("AUC:{}".format(mer15_el_result[4]))
print("PR_AUC:{}".format(mer15_el_result[5]))
```
```{python}
DF_k_sorted = DF_k.sort_values(by=["pep","HLA"]).reset_index(drop = True)
netMHCIIpan_ba_k = pd.read_csv("/Users/wangguangshuai/Data/MHCII/Data/IMM_model_benchmark/netMHCIIpan_k_mer_ba.csv")
netMHCIIpan_ba_k["pep_length"] =netMHCIIpan_ba_k["peptide"].map(len)
net_kmer_ba = netMHCIIpan_process(netMHCIIpan_ba_k,allele,DF_k_sorted,peptide)
kmer_ba_result = auc_pr(net_kmer_ba,"Immunogenicity","prediction",rank = True)
netMHCIIpan_el_k = pd.read_csv("../data/netMHCIIpan_k_mer_el.csv")
netMHCIIpan_el_k["pep_length"] =netMHCIIpan_el_k["peptide"].map(len)
net_kmer_el = netMHCIIpan_process(netMHCIIpan_el_k,allele,DF_k_sorted,peptide)
kmer_el_result = auc_pr(net_kmer_el,"Immunogenicity","prediction",rank = True)
print("BA AUC: {}".format(kmer_ba_result[4]))
print("BA PR_AUC: {}".format(kmer_ba_result[5]))
print("EL AUC: {}".format(kmer_el_result[4]))
print("EL PR_AUC: {}".format(kmer_el_result[5]))
```

```{python}
net_15mer_ba = net_15mer_ba.rename(columns = {"prediction":"net_prediction"})
p_test_15mer = pd.merge(result_15mer,net_15mer_ba)
z_score_15mer,p_val_15mer = DelongTest(p_test_15mer["prediction"],100-p_test_15mer["net_prediction"],p_test_15mer["Immunogenicity"]).show_result()
p_val_15mer
```

```{python}
net_kmer_ba = net_kmer_ba.rename(columns = {"prediction":"net_prediction"})
p_test_kmer = pd.merge(result_kmer,net_kmer_ba)
z_score_kmer,p_val_kmer = DelongTest(p_test_kmer["prediction"],100-p_test_kmer["net_prediction"],p_test_kmer["Immunogenicity"]).show_result()
p_val_kmer
```

```{python 15mer_ROC_fig}
#python
fig,ax = plt.subplots(figsize = (8,8))
lw = 2
ax.plot([0, 1], [0, 1], color='navy', lw=lw, linestyle='--')
ax.plot(mer15_result[0],mer15_result[1],label='TLimmuno2 : (AUC={0:.4f})'.format(mer15_result[4]), lw=lw)
ax.plot(IEDB_score[0],IEDB_score[1],label='IEDB : (AUC={0:.4f})'.format(IEDB_score[4]), lw=lw)
ax.plot(Repitope_score[0],Repitope_score[1],label='Repitope : (AUC={0:.4f})'.format(Repitope_score[4]), lw=lw)
ax.plot(mer15_ba_result[0],mer15_ba_result[1],label='NetMHCIIpan_ba : (AUC={0:.4f})'.format(mer15_ba_result[4]), lw=lw)
ax.plot(mer15_el_result[0],mer15_el_result[1],label='NetMHCIIpan_el : (AUC={0:.4f})'.format(mer15_el_result[4]), lw=lw)
ax.set_ylim([0,1.05])
ax.set_xlim([0,1])
ax.set_xlabel('False Positive Rate',fontsize = 14)
ax.set_ylabel('True Positive Rate',fontsize = 14)
ax.legend(loc="lower right",fontsize = 12)
plt.savefig("../figure/benchmark_15mer.pdf",dpi = 300,transparent=True)
plt.show()
```

```{python kmer_ROC_fig}
#python
fig,ax = plt.subplots(figsize = (8,8))
lw = 2
ax.plot([0, 1], [0, 1], color='navy', lw=lw, linestyle='--')
ax.plot(kmer_result[0],kmer_result[1],label='TLimmuno2 : (AUC={0:.4f})'.format(kmer_result[4]))
ax.plot(kmer_ba_result[0],kmer_ba_result[1],label='NetMHCIIpan_ba : (AUC={0:.4f})'.format(kmer_ba_result[4]))
ax.plot(kmer_el_result[0],kmer_el_result[1],label='NetMHCIIpan_el : (AUC={0:.4f})'.format(kmer_el_result[4]))
ax.set_ylim([0,1.05])
ax.set_xlim([0,1])
ax.set_xlabel('False Positive Rate',fontsize = 14)
ax.set_ylabel('True Positive Rate',fontsize = 14)
ax.legend(loc="lower right",fontsize = 12)
plt.savefig("../figure/benchmark_kmer.pdf",dpi = 300,transparent=True)
plt.show()
```

```{python 15mer_PR_fig}
#python
fig,ax = plt.subplots(figsize = (8,8))
lw = 2
ax.plot(mer15_result[2],mer15_result[3],label='TIMM2pred : (AUC={0:.4f})'.format(mer15_result[5]))
ax.plot(IEDB_score[2],IEDB_score[3],label='IEDB : (AUC={0:.4f})'.format(IEDB_score[5]))
ax.plot(Repitope_score[2],Repitope_score[3],label='Repitope : (AUC={0:.4f})'.format(Repitope_score[5]))
ax.plot(mer15_ba_result[2],mer15_ba_result[3],label='NetMHCIIpan_ba : (AUC={0:.4f})'.format(mer15_ba_result[5]))
ax.plot(mer15_el_result[2],mer15_el_result[3],label='NetMHCIIpan_el : (AUC={0:.4f})'.format(mer15_el_result[5]))

plt.xlim([0.0, 1.0])
plt.ylim([0.0, 1.05])
plt.xlabel('Recall',fontsize = 14)
plt.ylabel('Precision',fontsize = 14)
plt.legend(loc="upper right",fontsize = 12)
plt.savefig("../figure/15mer_PR.pdf",dpi = 300,transparent=True)
plt.show()
```

```{python kmer_PR_fig}
#python
fig,ax = plt.subplots(figsize = (8,8))
lw = 2
ax.plot(kmer_result[2],kmer_result[3],label='TIMM2pred : (AUC={0:.4f})'.format(kmer_result[5]))
ax.plot(kmer_ba_result[2],kmer_ba_result[3],label='NetMHCIIpan_ba : (AUC={0:.4f})'.format(kmer_ba_result[5]))
ax.plot(kmer_el_result[2],kmer_el_result[3],label='NetMHCIIpan_el : (AUC={0:.4f})'.format(kmer_el_result[5]))
plt.xlim([0.0, 1.0])
plt.ylim([0.0, 1.05])
plt.xlabel('Recall',fontsize = 14)
plt.ylabel('Precision',fontsize = 14)
plt.legend(loc="upper right",fontsize = 12)
plt.savefig("../figure/kmer_pr.pdf",dpi = 300,transparent=True)
plt.show()
```

### confusion matrix

```{python}
result_15mer["pre_label"] = result_15mer["prediction"].map(lambda x : 1 if x>0.5 else 0)
result_15mer_CM = confusion_matrix(result_15mer["Immunogenicity"],result_15mer["pre_label"])
result_15mer_CM = result_15mer_CM/len(result_15mer)
result_kmer["pre_label"] = result_kmer["prediction"].map(lambda x : 1 if x>0.5 else 0)
result_kmer_CM = confusion_matrix(result_kmer["Immunogenicity"],result_kmer["pre_label"])
result_kmer_CM = result_kmer_CM/len(result_kmer)
IEDB_result["pre_label"] = IEDB_result["score"].map(lambda x : 1 if x>50 else 0)
IEDB_result_CM = confusion_matrix(IEDB_result["Immunogenicity"],IEDB_result["pre_label"])
IEDB_result_CM = IEDB_result_CM/len(IEDB_result)
Repitope_result["pre_label"] = Repitope_result["ImmunogenicityScore"].map(lambda x : 1 if x>0.5 else 0)
Repitope_result_CM = confusion_matrix(Repitope_result["Immunogenicity"],Repitope_result["pre_label"])
Repitope_result_CM = Repitope_result_CM/len(Repitope_result)

def net_CM(data):
  data["pre_label"] = data["prediction"].map(lambda x : 1 if x<10 else 0)
  data_CM = confusion_matrix(data["Immunogenicity"],data["pre_label"])
  data_CM = data_CM/len(data)
  return data_CM


net_15mer_ba_CM = net_CM(net_15mer_ba)
net_15mer_el_CM = net_CM(net_15mer_el)
net_kmer_ba_CM = net_CM(net_kmer_ba)
net_kmer_el_CM = net_CM(net_kmer_el)
CM_all = [result_15mer_CM,result_kmer_CM,IEDB_result_CM,Repitope_result_CM,net_15mer_ba_CM,net_15mer_el_CM,net_kmer_ba_CM,net_kmer_el_CM]
```

```{python}
CM_all = [result_15mer_CM,IEDB_result_CM,Repitope_result_CM,net_15mer_ba_CM,net_15mer_el_CM,result_kmer_CM,net_kmer_ba_CM,net_kmer_el_CM]
CM_label = ["TLimmuno2_15mer","IEDB","Repitope","NetMHCIIpanBA_15mer","NetMHCIIpanEL_15mer","TLimmuon2_kmer","NetMHCIIpanBA_kmer","NetMHCIIpanEL_kmer"]
def CM_plot(data,label):
  fig,ax = plt.subplots(figsize = (4,4))
  sns.heatmap(data,annot =True,cmap="YlGnBu",vmin=0, vmax=1,cbar = False)
  plt.ylabel("True label")
  plt.xlabel("Prediction label")
  plt.title("{}".format(label))
  
  plt.savefig("../figure/{}.pdf".format(label),dpi = 300,transparent=True)

for i,l in enumerate(CM_all):
  CM_plot(l,CM_label[i])
  
```
## another benchmark

### structure immportance

we compare TLIMM2perd, TLimmuno2_noba, and TLimmuno2_onlyba model structure, and we found TLimmuno2 is significant improment than others.

```{python}
kmer_no_ba = pd.read_csv("/Users/wangguangshuai/Data/MHCII/Data/IMM_model_benchmark/kmer_no_ba_result.csv")
sam_pep_ID = kmer_no_ba["pep_ID"].unique()
sam_pep_ID = kmer_no_ba["pep_ID"].unique()
res = []
p_ID = []
for a in sam_pep_ID:
    Mean = []
    ID = kmer_no_ba[kmer_no_ba["pep_ID"] == a]
    for i in ID["HLA"].unique():
        x = ID[ID["HLA"] == i]
        #x_mean = x["Rank"].mean()
        x_mean = x["prediction"].max()
        Mean.append(x_mean)
    res.append(max(Mean))
    p_ID.append(a)
dic = {"pep_ID":p_ID,"prediction":res}
kmer_no_ba_result = pd.DataFrame(dic)
kmer_no_ba_result = kmer_no_ba_result.merge(peptide,how = "inner",on = "pep_ID" )
kmer_no_ba = auc_pr(kmer_no_ba_result,"Immunogenicity","prediction")
print("AUC:{}".format(kmer_no_ba[4]))
print("PR_AUC:{}".format(kmer_no_ba[5]))
kmer_only_ba = pd.read_csv("/Users/wangguangshuai/Data/MHCII/Data/IMM_model_benchmark/kmer_onlyBA_result.csv")
sam_pep_ID = kmer_only_ba["pep_ID"].unique()
res = []
p_ID = []
for a in sam_pep_ID:
    Mean = []
    ID = kmer_only_ba[kmer_only_ba["pep_ID"] == a]
    for i in ID["HLA"].unique():
        x = ID[ID["HLA"] == i]
        #x_mean = x["Rank"].mean()
        x_mean = x["prediction"].max()
        Mean.append(x_mean)
    res.append(max(Mean))
    p_ID.append(a)
dic = {"pep_ID":p_ID,"prediction":res}
kmer_only_ba_result = pd.DataFrame(dic)
kmer_only_ba_result = kmer_only_ba_result.merge(peptide,how = "inner",on = "pep_ID" )
kmer_only_ba = auc_pr(kmer_only_ba_result,"Immunogenicity","prediction")
print("AUC:{}".format(kmer_only_ba[4]))
print("PR_AUC:{}".format(kmer_only_ba[5]))
z_score_noba,p_val_noba = DelongTest(result_kmer["prediction"],kmer_no_ba_result["prediction"],result_kmer["Immunogenicity"]).show_result()
z_score_only_ba,p_val_only_ba = DelongTest(result_kmer["prediction"],kmer_only_ba_result["prediction"],result_kmer["Immunogenicity"]).show_result()
fig,ax = plt.subplots(figsize = (8,8))
lw = 2
ax.plot([0, 1], [0, 1], color='navy', lw=lw, linestyle='--')
ax.plot(kmer_result[0],kmer_result[1],label='TLimmuno2 : (AUC={0:.4f})'.format(kmer_result[4]))
ax.plot(kmer_no_ba[0],kmer_no_ba[1],label='TLimmuno2_noba : (AUC={0:.4f})'.format(kmer_no_ba[4]))
ax.set_ylim([0,1.05])
ax.set_xlim([0,1])
ax.set_xlabel('False Positive Rate',fontsize = 14)
ax.set_ylabel('True Positive Rate',fontsize = 14)
ax.text(0.78,0.2,"P value: {:.4f}".format(p_val_noba),weight="bold")
ax.legend(loc="lower right",fontsize = 12)
plt.savefig("../figure/kmer_noba.pdf",dpi = 300,transparent=True)
plt.show()
fig,ax = plt.subplots(figsize = (8,8))
lw = 2
ax.plot([0, 1], [0, 1], color='navy', lw=lw, linestyle='--')
ax.plot(kmer_result[0],kmer_result[1],label='TLimmuno2 : (AUC={0:.4f})'.format(kmer_result[4]))
ax.plot(kmer_only_ba[0],kmer_only_ba[1],label='TLimmuno2_onlyba : (AUC={0:.4f})'.format(kmer_only_ba[4]))
ax.set_ylim([0,1.05])
ax.set_xlim([0,1])
ax.set_xlabel('False Positive Rate',fontsize = 14)
ax.set_ylabel('True Positive Rate',fontsize = 14)
ax.text(0.78,0.2,"P value: {:.4f}".format(p_val_only_ba),weight="bold")
ax.legend(loc="lower right",fontsize = 12)
plt.savefig("../figure/kmer_onlyba.pdf",dpi = 300,transparent=True)
plt.show()
```


### true neoantigen

```{python neoantigen}
#python
neoantigen = pd.read_csv("../data/neoantigen.csv")
pseudo = pd.read_table("../data/pseudosequence.2016.all.X.dat",header = None,names = ["allele","sequence"])
neoantigen_process = pd.merge(neoantigen,pseudo,left_on = "HLA",right_on = "allele").dropna()
neoantigen_process.to_csv("../data/neoantigen_pro.csv")
```

```{python neoantigen_res}
#python

result = pd.read_csv("../data/neoantigen_result.csv")
result = result[result["immunogenic"] == 1]
Rank = pd.melt(result[["WT_Rank","Mut_Rank"]],var_name = "Type",value_name = "Rank")
from scipy import stats
_,p = stats.wilcoxon(result["WT_Rank"],result["Mut_Rank"])
fig,ax = plt.subplots(figsize = (4,6))
ax = sns.boxplot(data = Rank,x = "Type",y = "Rank",width = 0.4,color = "skyblue",linewidth = 2,showfliers = False)
b = sns.stripplot(data = Rank,x = "Type",y = "Rank",color = "red")
x1, x2 = 0, 1   
y, h, col = Rank['Rank'].max() + 0.1, 0.1, 'k'
plt.plot([x1, x1, x2, x2], [y, y+h, y+h, y], lw=1, c=col)
plt.text((x1+x2)*.5, y+h, "**", ha='center', va='bottom', color=col)
ax.text(0.3,0.9,"P = {:.4f}".format(p),weight="bold")
plt.ylim([0,1.2])
plt.yticks(np.arange(0,1.2,0.2))
plt.xticks((0,1),labels = ["WT peptide","Neoantigen"])
plt.tight_layout()
sns.despine(offset = 3, trim = True)
plt.savefig("../figure/neoantigen.pdf",dpi = 300,transparent=True)
plt.show()
```

### position importance

#### run

```{python, eval = FALSE}
#python
#PBS
import pandas as pd
import tensorflow as tf
import sys
sys.path.append("/public/slst/home/wanggsh/Project/MHCII/utils")
from Blosum62 import blosum62
import numpy as np
import random
import matplotlib.pyplot as plt
import matplotlib as mpl
import os
#os.environ["CUDA_VISIBLE_DEVICES"]="-1" 


immuno_data = pd.read_csv("/public/slst/home/wanggsh/Data/MHCII_immuno/IMM_data_random2.csv")
immuno_data = immuno_data[immuno_data["label"] == 1]
immuno_data["length"] = immuno_data["Description"].map(len)
#choose 15mer
immuno_data = immuno_data[immuno_data["length"] == 15]
immuno_data = immuno_data.reset_index(drop=True)


immuno_data["pep_blosum"] = immuno_data["Description"].apply(blosum62,args=(21,))
immuno_data["MHC_blosum"] = immuno_data["sequence"].apply(blosum62,args=(34,))
peptide = np.empty((len(immuno_data),21,21))
for i in range(len(immuno_data)):
    peptide[i] = immuno_data["pep_blosum"][i].reshape((21,21))
#mhc 
MHC = np.empty((len(immuno_data),34,21))
for i in range(len(immuno_data)):
    MHC[i] = immuno_data["MHC_blosum"][i].reshape((34,21))
Ala = list(blosum62("A",1))

BAmodel = tf.keras.models.load_model("/public/slst/home/wanggsh/Project/Saved_model/new_BA_model/model1")
BAmodel = tf.keras.models.Model(inputs = BAmodel.input,outputs = BAmodel.layers[-2].output)
IMM = tf.keras.models.load_model("/public/slst/home/wanggsh/Project/Saved_model/MHCII_immuno/model")

n = 100
position_rank = np.empty([n,15])
for m in range(n):
    index = np.random.choice(np.arange(len(immuno_data)),2000)
    peptide_sam = peptide[index].copy()
    MHC_sam = MHC[index]
    BA = BAmodel.predict([peptide_sam,MHC_sam])
    pred_ori = IMM([peptide_sam,MHC_sam,BA]).numpy().mean()  
    array = []  
    for i in range(15):
        peptide_sam[:,i,:] = 0
        BA = BAmodel.predict([peptide_sam,MHC_sam])
        importance = pred_ori - IMM([peptide_sam,MHC_sam,BA]).numpy().mean()
        array.append(importance)
        peptide_sam = peptide[index].copy()
    ranking = np.argsort(array) + 1
    tmp = []
    for i in range(15):
        tmp.append(list(ranking).index(i+1))
    position_rank[m,:] = tmp
np.save("/public/slst/home/wanggsh/Data/MHCII_immuno/position_importance/position.npy",position_rank)

n = 100
position_rank_ala = np.empty([n,15])
for m in range(n):
    index = np.random.choice(np.arange(len(immuno_data)),2000)
    peptide_sam = peptide[index].copy()
    MHC_sam = MHC[index]
    BA = BAmodel.predict([peptide_sam,MHC_sam])
    pred_ori = IMM([peptide_sam,MHC_sam,BA]).numpy().mean()  
    array = []  
    for i in range(15):
        peptide_sam[:,i,:] = Ala
        BA = BAmodel.predict([peptide_sam,MHC_sam])
        importance = pred_ori - IMM([peptide_sam,MHC_sam,BA]).numpy().mean()
        array.append(importance)
        peptide_sam = peptide[index].copy()
    ranking = np.argsort(array) + 1
    tmp = []
    for i in range(15):
        tmp.append(list(ranking).index(i+1))
    position_rank[m,:] = tmp
np.save("/public/slst/home/wanggsh/Data/MHCII_immuno/position_importance/position_ala.npy",position_rank)
```
#### result

```{python, message = FALSE }
#python
from collections import Counter
position_rank = np.load("../data/position.npy")
cmap = mpl.cm.get_cmap('tab20')
delim = np.linspace(0,1,15)
colors = [mpl.colors.rgb2hex(cmap(i)[:4]) for i in delim]
fig,ax = plt.subplots(figsize = (6,4))
for i in np.arange(15):
    y = list(Counter(position_rank[:,i]+1).keys())
    s = list(Counter(position_rank[:,i]+1).values())
    ax.scatter([i for n in range(len(y))],y, s=[m*3.5 for m in s],c=colors[i])
ax.set_ylim(0,16)
ax.set_yticks(np.arange(15)+1)
ax.set_xticks(np.arange(15))
ax.set_xticklabels(['1','2','3','4','5','6','7','8','9','10','11','12','13','14','15'])
ax.set_xlabel('Zero setting')
ax.set_ylabel('Ranking(ascending)')
ax.grid(True,alpha=0.2)
h1 = [ax.plot([],[],color='grey',marker='o',markersize=i,ls='')[0] for i in range(8,15,2)]
leg1 = ax.legend(handles=h1,labels=[10,40,70,100],title='Frequency',loc='lower left',bbox_to_anchor=(0.8,0.6),frameon=False)
ax.add_artist(leg1)
plt.savefig("../figure/position_zero.pdf",dpi = 300,transparent=True)
plt.show()
```

```{python,message = FALSE}
#python
position_rank_ala = np.load("../data/position_ala.npy")
cmap = mpl.cm.get_cmap('tab20')
delim = np.linspace(0,1,15)
colors = [mpl.colors.rgb2hex(cmap(i)[:4]) for i in delim]
fig,ax = plt.subplots(figsize = (6,4))
for i in np.arange(15):
    y = list(Counter(position_rank_ala[:,i]+1).keys())
    s = list(Counter(position_rank_ala[:,i]+1).values())
    ax.scatter([i for n in range(len(y))],y, s=[m*3.5 for m in s],c=colors[i])
ax.set_ylim(0,16)
ax.set_yticks(np.arange(15)+1)
ax.set_xticks(np.arange(15))
ax.set_xticklabels(['1','2','3','4','5','6','7','8','9','10','11','12','13','14','15'])
ax.set_xlabel('Ala scanning')
ax.set_ylabel('Ranking(ascending)')
ax.grid(True,alpha=0.2)
h1 = [ax.plot([],[],color='grey',marker='o',markersize=i,ls='')[0] for i in range(8,15,2)]
leg1 = ax.legend(handles=h1,labels=[10,40,70,100],title='Frequency',loc='lower left',bbox_to_anchor=(0.8,0.6),frameon=False)
ax.add_artist(leg1)
plt.savefig("../figure/position_ala.pdf",dpi = 300,transparent=True)
plt.show()
```

### BA model

we also do the same things in BA model.

```{python,eval = FALSE}
import pandas as pd
import tensorflow as tf
import sys
sys.path.append("/public/slst/home/wanggsh/Project/MHCII/utils")
from Blosum62 import blosum62
import numpy as np
import random
import matplotlib.pyplot as plt
import matplotlib as mpl
import os
#os.environ["CUDA_VISIBLE_DEVICES"]="-1" 


immuno_data = pd.read_csv("/public/slst/home/wanggsh/Data/MHCII/netMHCpanII4.0_train_data/BA_data/BA_MHC_label.csv")
immuno_data = immuno_data[immuno_data["score"] == 1]
immuno_data["length"] = immuno_data["peptide"].map(len)
#choose 15mer
immuno_data = immuno_data[immuno_data["length"] == 15]
immuno_data = immuno_data.reset_index(drop=True)


immuno_data["pep_blosum"] = immuno_data["peptide"].apply(blosum62,args=(21,))
immuno_data["MHC_blosum"] = immuno_data["sequence"].apply(blosum62,args=(34,))
peptide = np.empty((len(immuno_data),21,21))
for i in range(len(immuno_data)):
    peptide[i] = immuno_data["pep_blosum"][i].reshape((21,21))
#mhc 
MHC = np.empty((len(immuno_data),34,21))
for i in range(len(immuno_data)):
    MHC[i] = immuno_data["MHC_blosum"][i].reshape((34,21))
Ala = list(blosum62("A",1))

BAmodel = tf.keras.models.load_model("/public/slst/home/wanggsh/Project/Saved_model/new_BA_model/model1")

n = 100
position_rank = np.empty([n,15])
for m in range(n):
    index = np.random.choice(np.arange(len(immuno_data)),2000)
    peptide_sam = peptide[index].copy()
    MHC_sam = MHC[index]
    pred_ori = BAmodel([peptide_sam,MHC_sam]).numpy().mean()  
    array = []  
    for i in range(15):
        peptide_sam[:,i,:] = 0
        importance = pred_ori - BAmodel([peptide_sam,MHC_sam]).numpy().mean()
        array.append(importance)
        peptide_sam = peptide[index].copy()
    ranking = np.argsort(array) + 1
    tmp = []
    for i in range(15):
        tmp.append(list(ranking).index(i+1))
    position_rank[m,:] = tmp
np.save("/public/slst/home/wanggsh/Data/MHCII_immuno/position_importance/position_zero_BA.npy",position_rank)

n = 100
position_rank_ala = np.empty([n,15])
for m in range(n):
    index = np.random.choice(np.arange(len(immuno_data)),2000)
    peptide_sam = peptide[index].copy()
    MHC_sam = MHC[index]
    pred_ori = BAmodel([peptide_sam,MHC_sam]).numpy().mean()  
    array = []  
    for i in range(15):
        peptide_sam[:,i,:] = Ala
        importance = pred_ori - BAmodel([peptide_sam,MHC_sam]).numpy().mean()
        array.append(importance)
        peptide_sam = peptide[index].copy()
    ranking = np.argsort(array) + 1
    tmp = []
    for i in range(15):
        tmp.append(list(ranking).index(i+1))
    position_rank[m,:] = tmp
np.save("/public/slst/home/wanggsh/Data/MHCII_immuno/position_importance/position_ala_BA.npy",position_rank)
```

```{python}
#python
from collections import Counter
position_rank = np.load("../data/position_zero_BA.npy")
cmap = mpl.cm.get_cmap('tab20')
delim = np.linspace(0,1,15)
colors = [mpl.colors.rgb2hex(cmap(i)[:4]) for i in delim]
fig,ax = plt.subplots(figsize = (6,6))
for i in np.arange(15):
    y = list(Counter(position_rank[:,i]+1).keys())
    s = list(Counter(position_rank[:,i]+1).values())
    ax.scatter([i for n in range(len(y))],y, s=[m*3.5 for m in s],c=colors[i])
ax.set_ylim(0,16)
ax.set_yticks(np.arange(15)+1)
ax.set_xticks(np.arange(15))
ax.set_xticklabels(['1','2','3','4','5','6','7','8','9','10','11','12','13','14','15'])
ax.set_xlabel('Zero setting')
ax.set_ylabel('Ranking(ascending)')
ax.grid(True,alpha=0.2)
h1 = [ax.plot([],[],color='grey',marker='o',markersize=i,ls='')[0] for i in range(8,15,2)]
leg1 = ax.legend(handles=h1,labels=[10,40,70,100],title='Frequency',loc='lower left',bbox_to_anchor=(0.98,0.2),frameon=False)
ax.add_artist(leg1)
plt.savefig("../figure/position_zero_BA.pdf",dpi = 300,bbox_inches='tight',transparent=True)
plt.show()
```

```{python}
#python
position_rank_ala = np.load("../data/position_ala_BA.npy")
cmap = mpl.cm.get_cmap('tab20')
delim = np.linspace(0,1,15)
colors = [mpl.colors.rgb2hex(cmap(i)[:4]) for i in delim]
fig,ax = plt.subplots(figsize = (6,6))
for i in np.arange(15):
    y = list(Counter(position_rank_ala[:,i]+1).keys())
    s = list(Counter(position_rank_ala[:,i]+1).values())
    ax.scatter([i for n in range(len(y))],y, s=[m*3.5 for m in s],c=colors[i])
ax.set_ylim(0,16)
ax.set_yticks(np.arange(15)+1)
ax.set_xticks(np.arange(15))
ax.set_xticklabels(['1','2','3','4','5','6','7','8','9','10','11','12','13','14','15'])
ax.set_xlabel('Ala scaning')
ax.set_ylabel('Ranking(ascending)')
ax.grid(True,alpha=0.2)
h1 = [ax.plot([],[],color='grey',marker='o',markersize=i,ls='')[0] for i in range(8,15,2)]
leg1 = ax.legend(handles=h1,labels=[10,40,70,100],title='Frequency',loc='lower left',bbox_to_anchor=(0.98,0.2),frameon=False)
ax.add_artist(leg1)
plt.savefig("../figure/position_ala_BA.pdf",dpi = 300,bbox_inches='tight',transparent=True)
plt.show()
```
